---
title: "Danceability Trends for Ed Sheeran, G.E.M, and JJ Lin on Spotify (2010-2024)"
subtitle: "Consistent Danceability Patterns for Ed Sheeran and GEM, JJ Lin Shows Fluctuations"
author: 
  - Xuanle Zhou
  - Dezhen Chen
  - Yongqi Liu
  - Ziyuan Shen
  - Leo Liu
thanks: "Code and data are available at: https://github.com/Cassieliu77/Spotify_Trend_Analysis.git "
date: today
date-format: long
abstract: "This study explores the evolution of danceability  in the music of Ed Sheeran, JJ Lin, and G.E.M. from 2010 to 2024. It reveals that Ed Sheeran consistently maintains a high level of danceability, while G.E.M. incorporates increasingly danceable elements over time. In contrast, JJ Lin exhibits significant fluctuations in his focus on danceability."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| echo: false
#| warning: false
library(knitr)
library(usethis)
library(devtools) 
library(tidyverse)
library(spotifyr)
library(ggplot2)

ed_sheeran <- readRDS(here::here("data/01-raw_data/ed_sheeran.rds"))
jj_lin <- readRDS(here::here("data/01-raw_data/jj_lin.rds"))
gem <- readRDS(here::here("data/01-raw_data/gem.rds"))
```


# Introduction

In recent years, the concept of danceability has emerged as a significant metric for evaluating the musical characteristics of various artists and genres. Danceability reflects how suitable a track is for dancing. A score of 0.0 indicates the least danceable track, while a score of 1.0 signifies the most danceable, as defined by Spotify @howlin2021patients. This paper examines the evolution of danceability in the music of three prominent artists—Ed Sheeran, JJ Lin, and G.E.M.—across their respective discographies over time. By analyzing data on their released albums from 2010 to 2024, this study aims to explore trends and patterns in how these artists have developed their sound concerning danceability. We find that the trend analysis of danceability over time reveals that Ed Sheeran maintains a consistently high level of danceability in his music, while G.E.M. shows moderate variability, gradually incorporating more danceable elements. In contrast, JJ Lin exhibits the most fluctuation, initially focusing less on danceability before recently shifting back toward more danceable music.

# Data {#sec-data}
@fig-dance shows the trend of danceability over time for each of the three artists Ed Sheeran, JJ Lin, and G.E.M. Ed Sheeran's music shows a relatively stable trend, with danceability consistently above 0.5. The data points are closely clustered, which suggests that Ed Sheeran remains consistent and prioritizes danceable elements, which may attract a mainstream audience that enjoys upbeat, rhythm-driven music.

G.E.M. also shows a relatively steady trend towards danceability but more variability than Ed Sheeran's music. G.E.M.'s trend line initially drops but then rises, which shows a gradual adaptation to incorporate more danceable elements in recent years. This trend suggests that G.E.M. has achieved a balance between maintaining her core musical style and skillfully evolving to meet audience expectations. Her danceability values vary between 0.4 and 0.75, indicating moderate musical exploration.

JJ Lin had the most variation in danceability, with fluctuations ranging from 0.3 to 0.7. His trend line shows a gradually decreasing trend before a slight increase in recent years, and this indicates a period of less focus on danceability and then a re-focus on creating danceable music. 

In contrast, Ed Sheeran tends to be in a stable mode, keeping his music highly danceable throughout his career. G.E.M. strikes a balance, showing some flexibility and moderate variation in danceability. JJ Lin has seen the most change, with a significant shift in the extent to which he emphasizes danceable elements in his songs. This comparison shows how each artist shapes the danceability of their music in different ways.

```{r}
#| label: fig-dance
#| fig-cap: Danceability of Ed sheeran, G.E.M, JJ Lin artisits from 2010 to 2024
#| echo: false
#| warning: false

rbind(ed_sheeran, gem, jj_lin) |>
  select(artist_name, album_release_date, danceability) |>
  mutate(album_release_date = ymd(album_release_date)) |>  # Convert album release date to date format
  ggplot(aes(x = album_release_date, y = danceability, color = artist_name)) +
  geom_point(alpha = 0.3) +  
  geom_smooth(se = FALSE) +  
  theme_minimal() +  
  facet_wrap(facets = vars(artist_name), dir = "v") +  
  labs(
    x = "Album Release Date",
    y = "Danceability",
    color = "Artist"
  ) +
  scale_color_brewer(palette = "Set1") + 
  scale_x_date(limits = c(ymd("2010-01-01"), ymd("2024-12-31")),  
               date_breaks = "5 years",  
               date_labels = "%Y") +  
  theme(legend.position = "bottom") 
```
\newpage

\appendix

# Appendix {-}
```{r}
#| label: tbl-summary1
#| tbl-cap: Summary Table for Ed Sheeran
#| echo: false
#| warning: false
#| message: false

ed_sheeran <- ed_sheeran %>%
  rename(`Artist Name`= artist_name, `Danceability` = danceability,`Duration` = duration_ms) %>%
  select(`Artist Name`, `Danceability`,`Duration`)
kable(head(ed_sheeran))
```

```{r}
#| label: tbl-summary2
#| tbl-cap: Summary Table for GEM
#| echo: false
#| warning: false
#| message: false

gem <- gem %>%
  rename(`Artist Name`= artist_name, `Danceability` = danceability,`Duration` = duration_ms) %>%
  select(`Artist Name`, `Danceability`,`Duration`)
kable(head(gem))
```

```{r}
#| label: tbl-summary3
#| tbl-cap: Summary Table for JJ Lin
#| echo: false
#| warning: false
#| message: false

jj_lin <- jj_lin %>%
  rename(`Artist Name`= artist_name, `Danceability` = danceability,`Duration` = duration_ms) %>%
  select(`Artist Name`, `Danceability`,`Duration`)
kable(head(jj_lin))
```

# Acknowledgement

The data used in this paper comes from @citespotifyr. This study uses R packages [@citeR] to clean and analyze the dataset, including libraries from spotifyr [@citespotifyr], tidyverse [@tidyverse], usethis [@usethis], devtools [@devtools], ggplot2[@citeggplot]  and knitr [@citeknitr]. The data we used was from Spotify [@citespotifyr].


\newpage


# References


